Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective

### 2025-05-06 MS2_Krill.Rmd
adding edits

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated catch-at-length model perspective"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Erica Jarvis Mason^[Antarctic División NOAA, USA]
  - Andrea Pinones ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Richard Methot ^[NOAA, Fishery Science División, NorthWeast, Seattle, USA]
output:
  bookdown::word_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
#csl: apa.csl
csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem management objectives in the Western Antarctic Peninsula (WAP). In this study, we implement an integrated Catch-at-Length model that not only allows the incorporation of diverse data sources related to krill population dynamics, but also enables the evaluation of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status. We explored multiple scenarios that accounted for ecosystem influences on krill population dynamics, yielding results that align with the expected behavior of a stock assessment framework tailored to this type of species. Performance analyses reveal systematic differences across scenarios, highlighting increased bias in fishing mortality (F) and spawning stock biomass (SSB) in models that exclude ecosystem variables. Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics. Despite ongoing challenges in incorporating ecosystem variables into integrated model like this, we argue that these factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through post hoc analyses. Overlooking these drivers may result in less accurate conclusions and limit the effectiveness of management recommendations. Integrated modeling approaches that draw on extensive biological, fishery, and environmental datasets provide a valuable foundation for developing management strategies aligned with CCAMLR’s conservation and sustainable use objectives.

*Keywords: Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, Stock Synthesis model, perfomance model, fisheries management, CCAMLR*

@@

Antarctic krill (*Euphausia superba*) is a keystone species in the Antarctic ecosystem and one of the largest populations on the planet in terms of biomass (REF). Currently, its population is predominantly concentrated in the Western Antarctic Peninsula (WAP), a critical feeding ground for predators and the primary area for krill harvesting over the past four decades (REF). At the same time, the WAP is one of the most vulnerable regions to climate change (CC) and has already undergone rapid environmental shifts [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a]. These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective. Given its ecological significance and commercial value, effective management of krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries.

Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has adopted a new management strategy (CCAMLR 2019) that follows an ecosystem-based approach. This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species. It includes three key components: (i) stock assessments to determine precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a risk assessment framework, now referred to as spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks.

As part of this strategy, CCAMLR has tested various modeling approaches to assess krill stock status and support fisheries management decisions [@Hill2016]. While model-based stock assessments have been widely used (REF), many struggle to align with management objectives or accurately represent krill population dynamics due to their spatial variability and ecosystem complexity. Traditionally, stock assessments have relied on single-source data, either from fisheries catches [@Kinzey2015a; @Wang2021] or population monitoring surveys [@Green2023; @Watters2013].  One widely applied approach is the General Yield Model (GRYM), which provides global catch limits based on predefined harvest thresholds [@Maschette2020]. However, such approaches often neglect key ecological drivers. Considering the influence of habitat on population dynamics, the role of environmental variability in shaping krill productivity, and the contribution of predators as a source of natural mortality, stock assessment frameworks should explicitly incorporate these ecosystem components to more accurately represent population processes.

To address this need, we propose a conceptual framework for implementing a statistical model that integrates multiple data sources to assess the population dynamics of Antarctic krill in the Western Antarctic Peninsula (WAP) over the past decades. This approach incorporates biological, fishery, survey, and ecosystem-related information to provide a more comprehensive understanding of krill stock fluctuations and productivity.

Specifically, we employ an integrated catch-at-length implemented in Stock Synthesis model [@Methot2013] that allows us to evaluate the effects of environmental drivers on recruitment and productivity, as well as the impact of predator components into an integrated stock assessment model to Antarctic krill. The primary objective of this study was to assess the statistical performance of different management metrics, population variables and parameters based on alternative approaches to incorporating both environmental and predator data into a typical krill stock assessment model. By analyzing differences in estimated biomass, recruitment patterns, and overall krill productivity across these scenarios, we aim to quantify the extent to which ecosystem factors drive krill population trends. This comparative approach enables us to assess the relative importance of environmental conditions and predator dynamics in shaping krill stocks, providing critical insights for both fisheries management and conservation efforts.

@@

The study area corresponds to Subarea 48.1 in the WAP, where most krill fishing activity occurs and where CCAMLR has been working to implement a new management strategy at a reduced spatial scale. To refine the spatial resolution of krill population dynamics and analysis, we adopted the five strata proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest [@Dornam2021] (Fig 1 S1). While these strata were originally designed for management purposes without considering specific biological or population characteristics, their subdivision enables us to analyze the data at a finer scale. This level of detail facilitates the identification of heterogeneity across various data sources, whether these originate from the fishery or from environmental survey data, providing a more nuanced understanding of krill population dynamics. This spatial segregation of information is incorporated into the stock assessment model, making it a spatially explicit model that accounts for regional variability in population dynamics and external factors.

## Conceptual Model

In this approach we include spatial structure how an implicit approach where different areas are treated as separate fleets [REF]. Differences among strata have been identified from various perspectives of krill population dynamics [REF], as well as from the influence of environmental variability within Subarea 48.1 [REF] (Fig 2 S1). This approach allows integrate spatial heterogeneity into the stock assessment model to ensure that spatially structured processes are adequately represented. Although strata are treated as separate fleets, recruitment and growth processes remain unified, ensuring that the overall krill stock is modeled as a single population while accounting for spatially structured impacts at the Subarea 48.1 level. Each stratum is assigned a unique fishery and survey history, reflecting differences in harvesting intensity and spatial exploitation patterns. Survey indices and length composition data are fleet-specific, allowing for variations in data collection methods and regional abundance estimates. Selectivity is modeled separately for each strata to account for potential variations in size composition due to spatial heterogeneity or environmental conditions. The model considers interannual fluctuations of growth dynamics, maturity and reproduction, with a spawning pulse per year occurring in January regarding @REF. This approach represents a spatially implicit model, as it captures spatial variability in exploitation patterns without explicitly modeling movement or recruitment connectivity between areas. This allows for a more realistic assessment of krill stock productivity, in line with CCAMLR’s ecosystem-based management framework.

@@

Multiple sources of data were incorporated into the krill population dynamics model to improve inference and parameter estimation. These data include records from the krill fishery monitoring program, such as standardized CPUE (kg/time haul) and length compositions, which have been consistently collected on board fishing vessels by scientific observers through the CCAMLR Scheme of International Scientific Observation (SISO). This dataset contains detailed information including vessel identity, nationality, and georeferenced positions used in the exploratory analysis. In addition, scientific research cruises conducted under the AMLR program provided critical data such as krill biomass indices by strata and biological characteristics of krill obtained via net sampling. Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) were formatted according to model evaluation template requirements (see Supplementary Material 2). Environmental and predator-related data were also incorporated, such as satellite-derived chlorophyll-a concentrations from the Copernicus Marine Service and predator abundance and diet composition, particularly from penguin colonies reported in the [MAPPPD database](https://www.penguinmap.com/mapppd/sources/). A summary of all data sources, including temporal coverage and spatial scale, is provided in Table \@ref(tab:data-sources).

@@

Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition in krill dynamics population was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit1). Source data and temporal scale used in this stock assessment is showed in Fig 4, S1.

```{r parainit1, eval=TRUE}
dir1.1<-here("s1.1") 
start1 <- SS_readstarter(file = file.path(dir1.1,
                                          "starter.ss"),
                              verbose = FALSE)
dat1 <- SS_readdat(file = file.path(dir1.1, start1$datfile),
                        verbose = FALSE)
ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
                                    start1$ctlfil),
                        verbose = FALSE,
                        use_datlist = TRUE, 
                   datlist = dat1)
fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
                                                "forecast.ss"),
                              verbose = FALSE)
parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
 row.names( parbio)<-c("Nat M",
                       "Lmin", 
                       "Lmax",
                       "VonBert K",
                       "CV young",
                       "CV old", 
                       "Wt a", 
                       "Wt b",
                       "L50%", 
                       "Mat slope")
 SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
 Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
 Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]

 parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
parInit %>% 
  kable(format = "latex",
        booktabs = TRUE, 
        caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_styling(latex_options = c("striped"), 
                full_width = FALSE, 
                font_size = 9) %>%
  pack_rows(index = c("Natural Mortality" = 1,
                      "Growth" = 5,
                      "Length-Weight Relation" = 2,
                      "Maturity" = 2,
                      "Stock-Recruit Relation" = 5,
                      "Catchability" = 2,
                      "Selectivity" = 4))
```

## Catch-to-Length Model Implementation

Like many invertebrates, krill present challenges for age determination [@Punt2013]. Historically, aging techniques have not been available for this species, although recent advancements have been made [@Kilada2017]. As a result, stock assessments have relied on catch-at-length models with age dynamics approximated through transition matrices, which estimate the probability of individuals belonging to specific age classes based on their length. Integrated models can effectively capture the age structure of such populations by transforming length observations into population-level dynamics. Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches. Length data are generally more abundant and cost-effective to collect [@Canales2021, @Chong2019], and, due to the correlation between length and age, they provide valuable information about population structure. To date, all integrated krill assessments have used catch-at-length frameworks [@Kinzey2015a; @Kinzey2019a; @Wang2021]. Stock assessment models often rely on Age-Length Keys (ALKs) to infer the age structure of a population from observed length-frequency data [@ICCA2003]. An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classes. ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases. This approach allows the transformation of catch-at-length data into age-structured information, which is essential for age-based modeling frameworks. The ALK matrix is typically defined as:

The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK. The majority of ALKs are based on growth parameters according to the most suitable formula for the species. In the case of krill, these parameters are derived from the von Bertalanffy growth curve. In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV. This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomass with model estimates of vulnerable biomass. Weight-at-length was assumed known and calculated using @Maschette2020 parameters in krill resulting in a AKL exposed in Figure \@ref(fig:AKL)).

```{r AKL, eval=TRUE, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in Subarea 48.1"}
knitr::include_graphics('Figs/AKL.png')
```

>>>>

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Erica Jarvis Mason^[Antarctic División NOAA, USA]
  - Andrea Pinones ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Richard Methot ^[NOAA, Fishery Science División, NorthWeast, Seattle, USA]
output:
  bookdown::word_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
#csl: apa.csl
csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem management objectives in the Western Antarctic Peninsula (WAP). In this study, we implement an integrated length-to-age model that not only allows the incorporation of diverse data sources related to krill population dynamics, but also enables the evaluation of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status. We explored multiple scenarios that accounted for ecosystem influences on krill population dynamics, yielding results that align with the expected behavior of a stock assessment framework tailored to this type of species. Performance analyses reveal systematic differences across scenarios, highlighting increased bias in fishing mortality (F) and spawning stock biomass (SSB) in models that exclude ecosystem variables. Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics. Despite ongoing challenges in incorporating ecosystem variables into integrated model like this, we argue that these factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through post hoc analyses. Overlooking these drivers may result in less accurate conclusions, hence limiting the effectiveness of management recommendations. Integrated modeling approaches that draw on extensive biological, fishery, and environmental datasets provide a valuable foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use.

*Keywords: Antarctic krill, population dynamics, length-to-age model, stock-recruitment, ecosystem variables, perfomance model, fisheries management, CCAMLR*

@@

Antarctic krill (*Euphausia superba*) is one of the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @Kawaguchi2024; @McBride2021]. Currently, the population, and consecuently, the fishery is predominantly concentrated in the Western Antarctic Peninsula (WAP). This area is also a critical feeding ground for predators such as penguins, seals, and whales, and the primary area for krill harvesting over the past four decades [@Gallagher2023a; @Watters2013; @Hill2016]. At the same time, the WAP is one of the regions most affected by climate change, experiencing rapid increases in ocean temperature, sea ice loss, and changes in primary productivity [@Lima2013; @Turner2005; @Carrasco2021; @Atkinson2019a; @Morley2020; @McBride2021]. These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective in such a complex and changing environment [@Warwick-Evans2022; @Kawaguchi2024].

Given its ecosystem significance and commercial value, effective management of krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries. Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) that follows an ecosystem-based approach has been developing [@Cavanagh2016; @Constable2023; @Constable2011; @Watters2013]. This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and time. It includes three key components: (i) stock assessments to determine precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a risk assessment framework, now referred to as spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks [@Kawaguchi2024; @Warwick-Evans2022].

To provide scientific advice and recommendations on the status and standing stock of krill, CCAMLR has historically relied on dedicated scientific surveys [@Watters2013; @Green2023; @Hill2019; @Dornam2021]. However, changes in survey design—both methodological and spatial—have raised concerns about the reliability of these estimates, and the high costs associated with such studies have further motivated the search for alternative stock estimation methods. More recently, and as part of a new strategy, CCAMLR has explored a variety of model-based (indirect) approaches to assess krill stock status and support fisheries management decisions [@Hill2019]. Some of these stock assessment models have used a single source of data—either fishery-dependent catches [@Wang2021] or survey-based monitoring programs [@Kinzey2011; @Kinzey2019a; @Green2023]. Nowadays, a model-based approach used for generating management recommendations in CCAMLR is the Generalized Yield Model (GRYM) [@Maschette2020], which allows adjustments to the decision rules to predefined captures set by CCAMLR [@Constable2000; @CCAMLR2021a; @CCAMLR2010]. However, these model-based approaches exhibit limitations, including an effective integration of multiple data sources, incorporation of ecosystem-level variables into population dynamics and/or failure to account for the high spatial and temporal variability intrinsic to krill systems [@Meyer2020; @Trathan2021].

Increasing evidence shows that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Flores2012; @Morley2020; @Saba2014; @Perry2020; @Veytia2021; @Walsh2020; @Murphy2007; @Vorrath2020]. Likewise, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Reisinger2022; @Riaz2023; @Baines2022; @Biuw2024; @Moosa2017]. There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions [REF]. Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Marshall2019; @Crone2019; @Richards2016]. Therefore, any stock assessment frameworks related with new strategy of CCAMLR should explicitly incorporate these ecosystem components to more accurately represent krill population processes.

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data consistency is often affected by changes in monitoring programs and sampling methodologies [REF], while some key regions of the Western Antarctic Peninsula (WAP) lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [REF]. Despite these challenges, substantial data exist on krill and their environment. Fishery monitoring programs provide detailed catch and effort data through onboard scientific observers. Ecosystem monitoring, like the "CCAMLR Ecosystem Monitoring Program" (CEMP), track predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.

This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations. Spatially explicit approaches include the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011]. For statistical catch-at-age assessments, key frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013] among others. These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and populations dynamics [@Hoyle2020].

This study aims to integrate different sources of available data on Antarctic krill (*Euphausia superba*) —including fishery operations (catches, effort distribution), biological characteristics (growth rates, reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (environmental forcing and predator-prey interactions)—into a unified stock assessment framework. Specifically, we employ an integrated length-to-age model implemented in Stock Synthesis (SS3), which enables the explicit inclusion of environmental covariates affecting  biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration. The primary objective is to evaluate the statistical performance of alternative model configurations representing different hypotheses about ecosystem influences using a suite of diagnostic tools and performance metrics commonly applied in contemporary stock assessment practices. Our results provide empirical evidence for the value of ecosystem-informed in model-based approach, demonstrating that the inclusion of environmental and predator covariates has impact in  model fit, predictive skill and structural uncertainty.

This approach seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach [@CCAMLRWGSAM2024; @Hill2019; @Constable2023; @Zhao2023; @WGEMM2024; @Cavanagh2016; @Constable2011]. Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.

@@

The study area covers the Western Antarctic Peninsula (WAP), corresponding to FAO Subarea 48.1, which hosts a high of krill fishing activity. This is also the region where CCAMLR is actively developing a revised management approach at a finer spatial scale. To enhance the spatial resolution of our analysis of krill population dynamics, we adopted the five Management Units (MUs) (strata) proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig. 1, S1). It is worth noting that these strata were originally developed for management purposes, without explicitly considering biological or population-specific criteria [@Zaldua2024a; @Constable2023; @Watters2013]. However, their subdivision enables a finer-scale analysis of krill data. This increased spatial resolution allows for the detection of heterogeneity among various data sources—whether from fisheries or environmental  variables—thus providing a more detailed understanding of krill population dynamics.

## Spatially Implicit Model

To represent spatial heterogeneity in krill dynamics we implemented a spatially implicit modeling approach based on the *"areas-as-fleets"* concept [@Punt2019; @Nielsen2021; @Waterhouse2014]. The areas-as-fleets approach aims treats distinct geographical regions as independent fleets, under the assumption that fishing and survey operations can exert different effects across space. In this approach, each strata within the Western Antarctic Peninsula is treated as a distinct fleet and allows a spatial differentiation in key parameters such as selectivity, catchability (*q*) and abundance indices. This is especially relevant for krill, given the spatial variability in both population processes and fishing effort, given the abundant evidence of spatial heterogeneity [@Perry2019; @Perry2020; @Reiss2020; @SantaCruz2022; @Veytia2021]. By modeling each MU as a separate fleet, we allow the assessment to estimate region-specific parameters that reflect this variability. Survey indices and biological compositions (e.g., length-frequency data) are assigned to their corresponding MU based on sampling location based in @Dornam2021. Similarly, fishery-dependent data—such as catch and composition—are stratified by MU according to the reported fishing effort allowing for regional variation in both abundance and observational processes. Given the difference en the population structure [@Kinzey2011; @Reiss2020; @Cutter2022], the selectivity is modeled separately for each strata to account for potential variations in size composition due to spatial heterogeneity or environmental conditions. The model considers interannual fluctuations of growth dynamics, maturity and reproduction, with a spawning pulse per year occurring in January-February [@Perry2019; @McBride2021]. While strata are modeled as separate fleets, the population dynamics—such as recruitment and growth—remain unified, representing the krill stock as a single, interconnected population. This disaggregated structure enables the model to account for differences in small scale spatial dynamics and providing a flexible yet informative way to include spatial processes in stock assessments. Areas-as-fleet approach captures regional variability in dynamic and observation while avoiding the complexity of modeling explicit movement. This approach provides a more realistic representation of krill productivity within FAO Subarea 48.1.

## Length-to-Age Model Implementation

Like many invertebrates, krill present challenges for age determination [@Punt2013]. Historically, aging techniques have not been available for this species, although recent advancements have been made [@Kilada2017]. As a result, stock assessments have relied on length-to-age models with age dynamics approximated through transition matrices, which estimate the probability of individuals belonging to specific age classes based on their length. Integrated models can effectively capture the age structure of such populations by transforming length observations into population-level dynamics. Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches. Length data are generally more abundant and cost-effective to collect [@Canales2021, @Chong2019], and, due to the correlation between length and age, they provide valuable information about population structure. To date, all integrated krill assessments have used length-to-age frameworks [@Kinzey2015a; @Kinzey2019a; @Wang2021]. Stock assessment models often rely on Age-Length Keys (ALKs) to infer the age structure of a population from observed length-frequency data [@ICCA2003]. An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classes [@Lee2024]. ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases. This approach allows the transformation of catch-at-length data into age-structured information, which is essential for age-based modeling frameworks. The ALK matrix is typically defined as:

The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK. The majority of ALKs are based on growth parameters according to the most suitable formula for the species. In the case of krill, these parameters are derived from the von Bertalanffy growth curve. In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV. This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomass with model estimates of vulnerable biomass. Weight-at-length was assumed known and calculated using @Maschette2020 parameters in krill resulting in a AKL exposed in S1 Fig 7.

Growth was modeled following a von Bertalanffy growth function with lenght-at-age parameterization. The model assumed linear growth below a predefined settlement age of 1.5 years. The asymptotic length ($L_{\infty}$) was determined by the estimated length-at-age, indicating that the model used $L_{\infty}$ as an upper growth boundary. Exponential decay of growth beyond maximum age was disabled, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints. Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age. Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years. Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$. The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan. The model used a direct parameterization without additional offsets between sexes. Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters. Regarding selectivity functions, it was modeled using a length-based logistic formulation for the fishery, survey, and predator fleets. The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width). The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width. The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm depending of strata. Parameter estimation phases were assigned, with negative values indicating fixed parameters. The survey selectivity was also modeled using a length-based logistic function. Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm. The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values). These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets (Fig 6, S1). The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm. assuming that the predator can capture a wider range of krill sizes. These parameters were estimated in later phases, allowing flexibility in their estimation. Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization in a initial step. The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies. This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@REF]. The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value of 23, with a prior of 10. The parameter was freely estimated in phase 1.  The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability. The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts. Recruitment deviations were estimated to account for interannual variability in recruitment dynamics. The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term. The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years. Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed. Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment. To prevent overparameterization, recruitment deviations were constrained between -5 and 5. Cyclic patterns in recruitment were assumed each tree and five years [@REF], and advanced recruitment options were enabled for model flexibility. All of these configurations can be reviewed in the repository section of S1.

@@

A set of multiple sources of data were incorporated into the krill population dynamics model to improve inference and parameter estimation. These data included register from the krill fishery monitoring program, such as standardized CPUE (kg/time haul) and length compositions, which have been systematically collected on board fishing vessels by scientific observers through the "Scheme of International Scientific Observation" (SISO) Program. Scientific research cruises conducted under the AMLR program provided critical data such as krill biomass indices by strata and biological characteristics of krill obtained via net sampling. Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) were handled according to stock assessment model templates requirements (Suppl. Mat. 2). Environmental and predator-related data were also incorporated, such as satellite-derived chlorophyll-a concentrations from the Copernicus Marine Service and predator abundance and diet composition, particularly from penguin colonies reported in the [MAPPPD database](https://www.penguinmap.com/mapppd/sources/). A summary of all data sources, including temporal coverage and spatial scale, is provided in Table \@ref(tab:data-sources).

@@

Life history parameters, like growth, weight-length relation, natural mortality and maturity were used as priors to model initial condition in krill dynamics population was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit1). Source data and temporal scale used in this stock assessment are shown in Fig 4, S1.

```{r parainit1, eval=TRUE}
dir1.1 <- here("s1.1")

start1 <- SS_readstarter(file = file.path(dir1.1, "starter.ss"), verbose = FALSE)
dat1   <- SS_readdat(file = file.path(dir1.1, start1$datfile), verbose = FALSE)
ctl1   <- SS_readctl(file = file.path(dir1.1, start1$ctlfil),
                     verbose = FALSE, use_datlist = TRUE, datlist = dat1)
fore1  <- SS_readforecast(file = file.path(dir1.1, "forecast.ss"), verbose = FALSE)

# Extraer parámetros
parbio <- ctl1$MG_parms[1:10, c(1:4,7)]
rownames(parbio) <- c("Nat M", "Lmin", "Lmax", "VonBert K",
                      "CV young", "CV old", "Wt a", "Wt b", "L50%", "Mat slope")

SRpar  <- ctl1$SR_parms[1:5, c(1:4,7)]
Qpar   <- ctl1$Q_parms[1:2, c(1:4,7)]
Selpar <- ctl1$size_selex_parms[1:20, c(1:4,7)]

# Combinar parámetros
parInit <- as.data.frame(rbind(parbio, SRpar, Qpar, Selpar))

# Añadir nombres de fila como columna "Parameter"
parInit$Parameter <- rownames(parInit)

# Añadir categorías y reordenar columnas
parInit <- parInit %>%
  mutate(INIT  = round(INIT, 2),
         PRIOR = round(PRIOR, 2))

# Crear vector de categorías (39 filas)
categories <- c(rep("Natural Mortality", 1),
                rep("Growth", 4),
                rep("Length-Weight Relation", 2),
                rep("Maturity", 2),
                rep("Stock-Recruit Relation", 5),
                rep("Catchability", 2),
                rep("Selectivity", 21))  # OJO: corregido a 23

# Añadir categorías y reordenar columnas
parInit <- parInit %>%
  mutate(Category = categories) %>%
  dplyr::relocate(Category, Parameter)

# Crear flextable
ft <- flextable(parInit)
ft <- set_caption(ft, "Input parameters for the initial SS3 krill model")
ft <- autofit(ft)
ft <- theme_booktabs(ft)
ft <- fontsize(ft, size = 9, part = "all")
ft <- add_header_lines(ft, values = "Each parameter includes limits (LO-HI), initial value (INIT), and phase. Negative phase means the parameter is fixed.")

# Mostrar flextable
ft
